Distributed Tic Tac Toe:\\ Peer-to-Peer Team-Based Gaming System

### 2017-11-23 paper/paper.tex
Added NAT-traversal comments to paper as Architecture Evaluation section

By considering the fact that only the two leaders communicate about the votes between the two teams, then the system can easily break if one of the leaders creashes. 
In order to avoid that and provide a more robust system we chose to implemet the Bully Algorithm on each of the players so they will be able to detect a leader crash if they do not receive a vote reply from them in a certain period of time.
After that, they can easily call for an election on older processes than them and wait for a leader to be announced on a different period of time.
If no leader is announced on that period of time or there are no older processes, then that process can declare itself leader to the rest of the team and the opponent leader.

@@

The implementation follows the RMI distributed object model. A GameInt interface extends java.rmi.Remote and lists remote classes. 
A Game class implements GameInt and extends the UnicastRemoteObject. It is an analogy for what a player knows and does during a game. 
It provides an implementation for all the listed classes in the interface which include the functionality of requesting the player to make a play, 
setting the leader of the game, several methods to interact with the Role of the player and a main method that connects to other game instances. 
Currently, when running the game, the user passes it a local instance name and the name of another instance as arguments. Each instance then uses this information 
to do an RMI Registry lookup and get a reference to the other client. The Role interface allows the application to differentiate between Leader and Player classes which extend it. 
It utilises the State design pattern to allow players to seamlessly have their role changed – for example when the current leader disconnects. 
A Leader is responsible of collecting the votes from all of the members in their team in the turnStarts() method. This is done asynchronously using
a separate thread for each member – an instance of VoteThread. Each thread collects a vote from one assigned player then writes it into a corresponding position
in a shared AtomicIntegerArray that they have all been passed by the Leader. This ensures the main Leader thread can then read these values. 
Each Leader then informs all players in the game about the vote of his team. A Player instance represents a regular player that only must vote. 
Each regular player also has a mechanism to detect if a leader has not contacted them to choose a play for too long. If a time-out is triggered, 
a player calls for a new leader election. To avoid many simultaneous election calls, each regular player waits 5 minutes plus a randomised amount of time.
All plays are recorded on a GameState object instance that acts as a game board. After each turn it checks if the game is over and relays this message to the leaders
 which can then inform the members of their team.

@@

\section{Evaluation}

>>>>

By considering the fact that only the two leaders communicate about the votes between the two teams, then the system can easily break if one of the leaders crashes.

In order to avoid that and provide a more robust system we chose to implemet the Bully Algorithm on each of the players so they will be able to detect a leader crash if they do not receive a vote reply from them in a certain period of time.

After that, they can easily call for an election on older processes than them and wait for a leader to be announced on a different period of time.

If no leader is announced on that period of time or there are no older processes, then that process can declare itself leader to the rest of the team and the opponent leader.

@@

The implementation follows the RMI distributed object model. A GameInt interface extends java.rmi.Remote and lists remote classes.

A Game class implements GameInt and extends the UnicastRemoteObject. It is an analogy for what a player knows and does during a game. It provides an implementation for all the listed classes in the interface which include the functionality of requesting the player to make a play, setting the leader of the game, several methods to interact with the Role of the player and a main method that connects to other game instances. 
C
urrently, when running the game, the user passes it a local instance name and the name of another instance as arguments. Each instance then uses this information to do an RMI Registry lookup and get a reference to the other client. The Role interface allows the application to differentiate between Leader and Player classes which extend it.

It utilises the State design pattern to allow players to seamlessly have their role changed – for example when the current leader disconnects.

A Leader is responsible of collecting the votes from all of the members in their team in the turnStarts() method. This is done asynchronously using a separate thread for each member – an instance of VoteThread. Each thread collects a vote from one assigned player then writes it into a corresponding position in a shared AtomicIntegerArray that they have all been passed by the Leader. This ensures the main Leader thread can then read these values.

Each Leader then informs all players in the game about the vote of his team. A Player instance represents a regular player that only must vote.

Each regular player also has a mechanism to detect if a leader has not contacted them to choose a play for too long. If a time-out is triggered, a player calls for a new leader election. To avoid many simultaneous election calls, each regular player waits 5 minutes plus a randomised amount of time.

All plays are recorded on a GameState object instance that acts as a game board. After each turn it checks if the game is over and relays this message to the leaders which can then inform the members of their team.

@@

\section{Performance Evaluation}

\section{Architecture Evaluation}
While Java RMI provides a simplified approach to object manipulation over distributed systems, it does force the use of TCP sockets by default. The reliable, ordered delivery of data provided by TCP is desirable for distributed applications, however, this choice of transport protocol does have significant implications for application topology.

The ongoing shortage of IPv4 addresses has led to a large proportion of machines on the Internet to be obscured by Network Address Translation (NAT) middleboxes. Machines hidden behind one of these middleboxes cannot be contacted directly without the use of techniques like hole punching, which is a significant problem for establishing an entirely decentralised peer-to-peer system.

Hole punching for NAT traversal requires the use of an external, publicly-accessible proxy server to distribute the public-facing IP address and port combinations (addr:port) of NAT-obscured clients which want to establish a peer-to-peer session. This traversal method is used by many popular distributed applications such as multiplayer gaming, Skype, and peer-to-peer filesharing. This technique takes advantage of the behaviour of NAT middleboxes to create and maintain an 'open' addr:port combination - without this, a fully decentralised peer-to-peer session cannot be established.

As mentioned above, Java RMI uses TCP for its operations by default. While it is possible to modify the sockets created by RMI calls to use UDP instead, this requires extensive work using RMISocketFactory which is beyond the time allowed by the project.

UDP is the more commonly used transport protocol for peer-to-peer applications - one socket can be used for both sending and receiving messages from multiple clients without the need for establishing a fixed connection between two endpoints. Although adjustments would need to be made to allow for unordered or lost messages, UDP would allow us to implement a topology with good fault tolerance and Internet-wide access, as displayed in figure (INCLUDE IMAGE HERE).

An alternative approach is to attempt TCP hole punching instead. This requires creating four TCP sockets with bound to the same local addr:port combination (ie. with SO\_REUSEADDR enabled) to establish a connection with a peer through a proxy server, as described in section 4 of a study by Ford, et al (INCLUDE CITATION HERE). Not only is this a significant overhead, with four sockets and multiple ports required for each peer a client wants to make contact with, it is also more complex to perform than the UDP equivalent; our simplified Python-based tests could not establish contact between peers using the instructions described in the Ford paper.

Without significant adjustments to the type of sockets used by Java RMI, a better understanding of TCP hole punching, or the universal adoption of IPv6, our application is currently limited to operating within local networks only. This is acceptable for its use as an educational tool, where it can be used within corporate or university networks freely, but the lack of NAT traversal largely excludes its use as a casual multiplayer game given that the vast majority of homes obscure several devices behind an ISP-provided NAT middlebox.

While we are aiming to create a system without unnecessary complexity for educational purposes, NAT is a core component of the modern Internet - all networks programmers will encounter difficulties with it at some stage and having a solution to demonstrate as part of our application would have been very beneficial to people using it for educational purposes.
